---
title: "Intro to Raster Data in R"
author: "Data Carpentry"
date: "Started 2020-03-26"
output:
  knitr:::html_vignette:
    css: inc/jm-gray-vignette.css
    number_sections: yes
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment=NA)
```

This document originated at [datacarpentry.org](https://datacarpentry.org/r-raster-vector-geospatial/01-raster-structure/index.html).

# Introduction

## Questions

- What is a raster dataset?

- How do I work with and plot raster data in R?

- How can I handle missing or bad data values for a raster?

## Objectives

- Describe the fundamental attributes of a raster dataset.

- Explore raster attributes and metadata using R.

- Import rasters into R using the raster package.

- Plot a raster file in R using the `ggplot2` package.

- Describe the difference between single- and multi-band rasters.

## Things needed to complete this task

See the [lesson homepage](https://datacarpentry.org/r-raster-vector-geospatial/) for detailed information about the software, data, and other prerequisites you will need to work through the examples in this episode.

In this episode, we will introduce the fundamental principles, packages
and metadata/raster attributes that are needed to work with raster data
in R. We will discuss some of the core metadata elements that we need
to understand to work with rasters in R, including CRS and resolution.
We will also explore missing and bad data values as stored in a raster
and how R handles these elements.

We will continue to work with the `dplyr` and `ggplot2` packages that were introduced in the
[Introduction to R for Geospatial Data](https://datacarpentry.org/r-raster-vector-geospatial/)
lesson. We will use two additional packages in this episode to work with raster data - the `raster` and `rgdal` packages. Make sure that you have these packages loaded.

```{r loadPackages, comment=NA}
library(dplyr)
library(raster)
library(rgdal)
library(here)
library(ggplot2)
```

## Introduce the Data

If not already discussed, introduce the datasets that will be used in this lesson. A brief introduction to the datasets can be found on the [Geospatial workshop homepage](https://datacarpentry.org/geospatial-workshop/).

For more detailed information about the datasets, check out the [Geospatial workshop data page](http://datacarpentry.org/geospatial-workshop/data/).

# View Raster File Attributes

We will be working with a series of GeoTIFF files in this lesson. The
GeoTIFF format contains a set of embedded tags with metadata about the
raster data. We can use the function `GDALinfo()` to get information
about our raster data before we read that data into R. It is ideal to
do this before importing your data.

```{r checkData}
fi <- paste0(here(),"/data/tif/HARV_dsmCrop.tif")
GDALinfo(fi)
```

```{r captureOutput}
HARV_dsmCrop_info <- capture.output(GDALinfo(fi))
```

# Open a raster in R


```{r loadRaster}
DSM_HARV <- raster(fi)
DSM_HARV
```

The information above includes a report of min and max values, but no
other data range statistics. Similar to other R data structures like
vectors and data frame columns, descriptive statistics for raster data
can be retrieved like

```{r getSummary}
summary(DSM_HARV)
```

Note the warning - unless you force R to calculate these statistics
using every cell in the raster, it will take a random sample of
100,000 cells and calculate from that instead. To force calculation
on more, or even all values, you can use the parameter `maxsamp`:

```{r limitedSummary}
summary(DSM_HARV, maxsamp = ncell(DSM_HARV))
```

You may not see major differences in summary stats as `maxsamp` increases, except with very large rasters.

To visualize this data in R using `ggplot2`, we need to convert it
to a `dataframe`. We learned about dataframes in an earlier lesson.
The `raster` package has an built-in function for conversion to a
plotable dataframe.

```{r makeDF}
DSM_HARV_df <- as.data.frame(DSM_HARV, xy = TRUE)
```

Now when we view the structure of our data, we will see a standard
`dataframe` format.

```{r getStructure}
str(DSM_HARV_df)
```

We can use `ggplot()` to plot this data. We will set the color scale
to `scale_fill_viridis_c` which is a color-blindness friendly color
scale. We will also use the `coord_quickmap()` function to use an
approximate Mercator projection for our plots. This approximation is
suitable for small areas that are not too close to the poles. Other
coordinate systems are available in `ggplot2` if needed, you can learn
about them at their help page `?coord_map`.

```{r makeFirstPlot}
ggplot() +
    geom_raster(data = DSM_HARV_df ,
                aes(x = x, y = y,
                fill = HARV_dsmCrop)) +
    scale_fill_viridis_c() +
    coord_quickmap()

```

# More Info

- More information about the Viridis palette used above at R Viridis package [documentation](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html).

- For faster, simpler plots, you can use the `plot` function
from the `raster` package.

